# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data

```{r echo=TRUE}
activity <- read.csv(file = "activity.csv", na.strings = "NA")
activityCompleteCases <- activity[complete.cases(activity),]
activity$date <- as.Date(activity$date)
```

## What is mean total number of steps taken per day?

```{r, echo=TRUE, fig.height=4, fig.width=4}
# Calculate total number of steps per day
stepsPerDay <- aggregate(steps ~ date, activityCompleteCases, sum)
# Create histogram of total nummber of steps per day
hist(stepsPerDay$steps, 
     main="Distribution of total steps taken/day", 
     xlab = "Total # of Steps taken", 
     ylab = "# of Days")
meanStepsPerDay <- round(mean(stepsPerDay$steps), 0)
medianStepsPerDay <- round(median(stepsPerDay$steps) ,0)
```

* Mean steps per day: `r as.integer(meanStepsPerDay) `
* Median steps per day: `r as.integer(medianStepsPerDay) `

## Average daily activity pattern
```{r, echo=TRUE, fig.height=4, fig.width=4}
# Average number of steps by 5-minute interval
avgStepsPerInterval = aggregate(steps ~ interval, activityCompleteCases, mean)
# Time-series plot of average number of steps by 5-minute interval
plot(avgStepsPerInterval$interval, avgStepsPerInterval$steps, type = "l", 
     main = "Average steps per interval",
     xlab = "Interval",
     ylab = "Average Steps")
intervalWithMostSteps <- avgStepsPerInterval[avgStepsPerInterval$steps == max(avgStepsPerInterval$steps),]$interval
```
The 5-minute interval that, on average, contains the maximum number of steps is interval `r intervalWithMostSteps `.

## Imputing missing values

```{r, echo=TRUE, fig.height=4, fig.width=4}
# Number of incomplete cases
countIncompleteCases <- sum(!complete.cases(activity))
# Fill missing values with mean for given 5-minute interval
activityImputed <- merge(activity, avgStepsPerInterval, by = "interval")
colnames(activityImputed) <- c("interval", "steps", "date", "imputedSteps")
activityImputed$steps[is.na(activityImputed$steps)] <- activityImputed$imputedSteps[is.na(activityImputed$steps)]
activityImputed <- activityImputed[order(activityImputed$date, activityImputed$interval),]

# Calculate total number of steps per day
imputedStepsPerDay <- aggregate(steps ~ date, activityImputed, sum)

# Create histogram of imputed total nummber of steps per day
hist(imputedStepsPerDay$steps, 
     main="Imputed Mean Steps Per Day", 
     xlab = "Average Steps", 
     ylab = "# of Days")
# Calculate and report mean and median
meanStepsPerDayImputed <- round(mean(imputedStepsPerDay$steps), 0)
medianStepsPerDayImputed <- round(median(imputedStepsPerDay$steps), 0)
```
* Imputed mean steps per day: `r as.integer(meanStepsPerDayImputed) `
* Imputed median steps per day: `r as.integer(medianStepsPerDayImputed) `

## Are there differences in activity patterns between weekdays and weekends?
```{r, echo=TRUE, fig.height=8, fig.width=4}
activityImputed$partOfWeek <- factor (
  ifelse(is.element(weekdays(activityImputed$date), c("Saturday", "Sunday")),
         "weekend", "weekday"))

# Average number of steps by 5-minute interval, for weekdays and weekends
imputedAvgStepsPerIntervalWeekday = aggregate(steps ~ interval, activityImputed[activityImputed$partOfWeek == "weekday",], mean)
imputedAvgStepsPerIntervalWeekend = aggregate(steps ~ interval, activityImputed[activityImputed$partOfWeek == "weekend",], mean)

# Create panel plot by weekday vs. weekend
par(mfrow=c(2,1))

# Time-series plot of average number of steps by 5-minute interval, weekend
plot(imputedAvgStepsPerIntervalWeekend$interval, 
     imputedAvgStepsPerIntervalWeekend$steps, 
     type = "l", 
     main = "Weekend",
     xlab = "5-minute interval",
     ylab = "Average # of steps")

# Time-series plot of average number of steps by 5-minute interval, weekday
plot(imputedAvgStepsPerIntervalWeekday$interval, 
     imputedAvgStepsPerIntervalWeekday$steps, 
     type = "l", 
     main = "Weekday",
     xlab = "5-minute interval",
     ylab = "Average # of steps")
```
